\documentclass[11pt,letterpaper]{article}

\input{../../../../.config/latex/preamble_v1.tex}
\lightmode

\title{\textbf{Math 137 Problem Set 3}}

\begin{document}
\maketitle

Throughout, $K$ is assumed to be an algebraically closed field.

\begin{problem}
    Let $A$ be an irreducible algebraic subset of $K^n$ and let $B$ be an irreducible algebraic subset of $K^m$. Show that the cartesian product $A\times B$ is an irreducible algebraic subset of $K^n \times K^m = K^{n+m}$.
\end{problem}

\begin{problem}
    If $S$ is an integral ring extension of $R$ and $T$ is an integral ring extension of $S$, prove that $T$ is an integral ring extension of $R$.
\end{problem}

\begin{problem}
    Show that any integral ring-finite ring extension is module-finite.
\end{problem}

\begin{problem}
    Find a ring-finite ring extension $S$ of a ring $R$ for which $S$ is a field, but which is not module-finite.
\end{problem}

\begin{problem}\noindent
    \begin{enumerate}[(a)]
        \item What is the integral closure of $K[X^2,X^3]\subseteq K[X]$ in $K(X)$?
        \item Let $R=\mathbb Z/4\mathbb Z$. What is the integral closure of $R$ in $R[X]$?
    \end{enumerate}
\end{problem}

\begin{problem}
    Let $X$ be a topological space. Let $A\neq\emptyset$ be a closed subset and equip it with the subspace topology. Show that the following are equivalent:
    \begin{enumerate}[(a)]
        \item We cannot write $A=A_1\cup A_2$ with closed subsets $A_1,A_2\subsetneq A$.
        \item Any two open subsets $\emptyset\neq U_1,U_2\subseteq A$ (for the subspace topology) intersect.
        \item Every open subset $\emptyset\neq U\subseteq A$ (for the subspace topology) is dense in~$A$.
    \end{enumerate}
    Such a set $A$ is called \emph{irreducible}. (Note that the irreducible algebraic subsets of $K^n$ are exactly the irreducible closed subsets for the Zariski topology.)
\end{problem}

\begin{problem}
    Let $f:X\rightarrow Y$ be a continuous map between topological spaces. Let $A\subseteq X$ be a closed irreducible subset. Show that $\overline{f(A)}\subseteq Y$ is irreducible.
\end{problem}

\end{document}
