\documentclass{../../templates/lkx_pset}

\usepackage[T1]{fontenc}
\RequirePackage{mlmodern}

\title{Math 230a Problem Set 3}
\author{Lev Kruglyak}
\due{September 25, 2024}

\input{../math230a.sty}

\collaborator{AJ LaMotta}
\collaborator{Leonardo Kaplan}
\collaborator{Ignasi Vicente}

\begin{document}
\maketitle

\begin{problem}{1}
Distributions of rank $2$ on $3$-manifolds.
\end{problem}

\begin{parts}
	\begin{part}{a}
		Let $M$ be a $3$-manifold and $\alpha$ a non-zero $1$-form. Prove that the $2$-dimensional distribution determined by $\alpha$ is integrable if and only if $\alpha\wedge d\alpha = 0$.
	\end{part}

	Suppose $\alpha$ is a nowhere zero $1$-form on $M$. We can consider this a section $\alpha\in\Gamma(T^*M)$. Since $\alpha$ is nowhere zero, the kernel bundle $\ker\alpha$ is a $2$-dimensional subbundle of $TM$. Let's call this distribution $E_\alpha$. Clearly, a vector field $\xi\in \Gamma(TM)$ belongs to $E_\alpha$ if and only if $\alpha(\xi) = 0$. The distribution is integrable if and only if for all vector fields $\xi_1,\xi_2$ belonging to $E_\alpha$, their commutator $[\xi_1,\xi_2]$ also belongs to $E_\alpha$. This means that we can we must show that:
	\[
		\left\{\alpha\wedge d\alpha = 0\right\}  \quad\iff\quad \left\{
		\alpha([\xi_1, \xi_2]) =0
		\quad\textrm{whenever}\quad\alpha(\xi_1) = \alpha(\xi_2)=0\right\}.
	\]
	By the standard commutator relations between $d,\iota,\mathcal{L}$, for any vector fields $\xi_1$, $\xi_2$ belonging to $E_\alpha$, we have
	\[
		\begin{aligned}
			\alpha([\xi_1, \xi_2])
			 & = \iota_{[\xi_1, \xi_2]}\alpha                                                    \\
			 & = \mathcal{L}_{\xi_1}\iota_{\xi_2}\alpha + \iota_{\xi_2}\mathcal{L}_{\xi_1}\alpha \\
			 & = \iota_{\xi_2}\mathcal{L}_{\xi_1}\alpha                                          \\
			 & = \iota_{\xi_2}(d\iota_{\xi_1} + \iota_{\xi_1} d)\alpha                           \\
			 & = \iota_{\xi_2}d\iota_{\xi_1}\alpha + \iota_{\xi_1} d\alpha                       \\
			 & = \iota_{\xi_2}\iota_{\xi_1} d\alpha                                              \\
			 & = d\alpha(\xi_2, \xi_1)
		\end{aligned}
	\]
	Next, suppose $\xi_1,\xi_2,\xi_3$ are any vector fields, not necessarily in $E_\alpha$. Then, we have
	\[
		\begin{aligned}
			(\alpha\wedge d\alpha)(\xi_1, \xi_2, \xi_3)
			 & =\iota_{\xi_1}\iota_{\xi_2}\iota_{\xi_3} (\alpha\wedge d\alpha)                                                                                                                           \\
			 & =\iota_{\xi_1}\iota_{\xi_2} ( \iota_{\xi_3} \alpha\wedge d\alpha - \alpha\wedge \iota_{\xi_3} d\alpha)                                                                                    \\
			 & = \iota_{\xi_1}(-\iota_{\xi_3}\alpha \wedge \iota_{\xi_2}d\alpha - \iota_{\xi_2}\alpha\wedge \iota_{\xi_3} d\alpha + \alpha\wedge \iota_{\xi_2}\iota_{\xi_3}d\alpha)                      \\
			 & =\iota_{\xi_3}\alpha\wedge \iota_{\xi_1}\iota_{\xi_2} d\alpha + \iota_{\xi_2}\alpha\wedge \iota_{\xi_1}\iota_{\xi_3}d\alpha + \iota_{\xi_1}\alpha\wedge \iota_{\xi_2}\iota_{\xi_3}d\alpha \\
			 & = \alpha(\xi_3)\cdot d\alpha(\xi_1, \xi_2) + \alpha(\xi_2)\cdot d\alpha(\xi_1, \xi_3) + \alpha(\xi_1)\cdot d\alpha(\xi_2, \xi_3).
		\end{aligned}
	\]
	Here, we make use of the identity $\iota_{\xi}(\alpha\wedge \beta) = \iota_{\xi}\alpha\wedge \beta + (-1)^{|\alpha|}\alpha\wedge \iota_{\xi}\beta$, and cancel terms such as $\iota_{\xi_i}\iota_{\xi_j} \alpha$ and $\iota_{\xi_i}\iota_{\xi_j}\iota_{\xi_k}d\alpha$ since all negative degree forms are zero. With the two identities we derived, we can now get our result.

	First, suppose that $\alpha\wedge d\alpha=0$. Whenever we have vector fields $\xi_1, \xi_2$ belonging to $E_\alpha$, and a vector field $\eta$ not necessarily belonging to $E_\alpha$, note that
	\[
		\begin{aligned}
			0 = (\alpha\wedge d\alpha)(\xi_1, \xi_2, \eta)
			 & = \alpha(\eta)\cdot d\alpha(\xi_1, \xi_2) + \alpha(\xi_2)\cdot d\alpha(\xi_1, \eta) + \alpha(\xi_1)\cdot d\alpha(\xi_2, \eta) \\
			 & = \alpha(\eta)\cdot \alpha([\xi_2, \xi_1]).
		\end{aligned}
	\]
	Since $\alpha$ is nonzero, we can find some vector field $\eta$ such that $\alpha(\eta)$ is nonzero. This means that $\alpha([\xi_2,\xi_1])=0$, and so the distribution is integrable.

	In the converse direction, suppose that for any vector fields $\xi_1,\xi_2$ belonging to $E_\alpha$, we have $\alpha([\xi_1,\xi_2])=0$. Let $U$ be an open set on which we can find such fields $\xi_1, \xi_2 \in \Gamma(E_\alpha; U)$ which are also linearly independent. Finally, suppose that there is a third field $\eta \in \Gamma(TM; U)$, which is linearly independent to $\xi_1, \xi_2$ so that these fields span the tangent bundle $TM$ restricted to $U$. On $U$, we have
	\[
		(\alpha\wedge d\alpha)(\xi_1, \xi_2,\eta) = \alpha(\eta)\cdot \alpha([\xi_2, \xi_1]) = 0.
	\]
	However, these fields formed a local frame the tangent bundle, so it follows that $\alpha\wedge d\alpha$ vanishes on $U$. Since our choice of $U$ was arbitrary, we can do this over the entire manifold to show that $\alpha\wedge d\alpha$ vanishes globally.

	\begin{part}{b}
		The Hopf fibration $\pi : S^3 \to S^2$ may be constructed by identifying $S^3$ as the unit sphere in $\C^2$ and $S^2$ as $\CP^1$; then the map is a restriction of the canonical projection $(\C^2)^\times \to \CP^1$. The kernel $E' = \ker d\pi$ is an (integrable) one-dimensional distribution on $S^3$. Let $E\subset TS^3$ be the $2$-dimensional distribution given by the orthogonal complement of $E'$ with respect to the standard round metric. Is $E$ integrable? Find a nonzero $1$-form $\alpha$ which generates the ideal $\mathcal{I}(E)$ associated to $E$. Compute $d\alpha$ and $\alpha\wedge d\alpha$.
	\end{part}

	We can consider $S^3$ as the subset of $\C^2$ given by points $(z_1, z_2)$ with $|z_1|^2 + |z_2|^2=1$. Choosing coordinates:
	\[
		\begin{cases} z_1 = \cos(\theta)\cdot e^{i(\phi + \psi)} \\ z_2 = \sin(\theta)\cdot e^{i(\phi - \psi)} \end{cases} \qtq{where} (\theta,\phi,\psi)\in [0,\pi/2]\times[0,\pi]^2.
	\]
	In these coordinates, the Hopf fibration becomes $\pi(\theta,\phi,\psi) = (\theta,\phi)$ where we use the standard spherical coordinate system on $S^2$.

	Next, let's see what form the round metric takes on $S^3$ using the coordinate system we provided. Recall that the round metric on $S^3$ is the pullback of the Euclidean metric on $\R^4$ from the canonical embedding $\C^2\to \R^4$. Expanding our coordinate system in $\R^4$, and
	% Letting $x,y,z,h$ be the standard coordinates on $\R^4$, we see that
	% \[
	%     \begin{cases}
	%       x = \cos(\theta)\cdot \cos(\phi+\psi)\\
	%       y = \cos(\theta)\cdot \sin(\phi +\psi)\\
	%       z = \sin(\theta)\cdot \cos(\phi - \psi)\\
	%       h = \sin(\theta)\cdot \sin(\phi - \psi)
	%     \end{cases}\quad\implies\quad
	%     \begin{cases}
	%       dx = -\cos(\theta)\sin(\phi+\psi)(d\phi + d\psi)-\sin(\theta) \cos(\phi+\psi) \,d\theta\\
	%       dy = \cos(\theta)\cos(\phi+\psi)(d\phi + d\psi)-\sin(\theta) \sin(\phi+\psi) \,d\theta\\
	%       dz = -\sin(\theta)\sin(\phi+\psi)(d\phi + d\psi)+\cos(\theta) \cos(\phi+\psi) \,d\theta\\
	%       dh = \sin(\theta)\cos(\phi+\psi)(d\phi + d\psi)+\cos(\theta) \sin(\phi+\psi) \,d\theta\\
	%     \end{cases}
	% \]
	using standard trigonometric identities, up to scaling, the metric $g$ on $S^3$ has matrix form
	\[
		g =
		\begin{pmatrix} 1 & 0              & 0              \\
                0 & 1              & -\cos(2\theta) \\
                0 & -\cos(2\theta) & 1\end{pmatrix}\quad\iff\quad
		g = d\theta^2 + d\phi^2- 2\cos (2\theta) \,d\phi\, d\psi + d\psi^2.
	\]

	Now suppose at a point $(\theta,\phi,\psi)\in S^3$, the tangent vector $v\in T_{(\theta,\phi,\psi)} S^3$ is in the kernel of $d\pi$. This means that $v$ must have no $\partial/\partial \phi$ or $\partial /\partial \psi$ components. The space of vectors $v=(x,y,z)$ in the complement of this kernel must then satisfy, for every $t\in \R$,
	\[
		\begin{pmatrix}0 \\ 0 \\ t\end{pmatrix}^\intercal
		\begin{pmatrix} 1 & 0              & 0              \\
                0 & 1              & -\cos(2\theta) \\
                0 & -\cos(2\theta) & 1\end{pmatrix}\begin{pmatrix}x\\y\\z\end{pmatrix} = 0\quad\implies\quad -\cos (2\theta) y + z = 0.
	\]
	In other words, $E$ is spanned by the vector fields
	\[
		\xi_1= \frac{\partial}{\partial \theta}
		\quad\textrm{ and }\quad
		\xi_2 = \frac{\partial}{\partial \psi} - \cos(2\theta)\frac{\partial}{\partial \phi}.
	\]
	This distribution is thus \textbf{not integrable}, since a simple calculation shows that the commutator of these vector fields is $[\xi_1, \xi_2] = 2\sin(2\theta)\partial / \partial \phi$, which does not belong to $E$.

	Next, we need to find a 1-form $\alpha$ which generated the ideal associated to the distribution. Since we have vector fields which already span $E$, it's clear that the $1$-form $\alpha = \cos(2\theta)\,d\psi + d\phi$ vanishes on $E$, and clearly must be a generator for $\mathcal{I}(E)$ since it has minimal degree. Computing $d\alpha$ and $\alpha\wedge d\alpha$, we get
	\[
		d\alpha = -2\sin(2\theta)\,d\theta\wedge d\psi \quad\implies\quad \alpha\wedge d\alpha = (2\cos(2\theta)d\psi - d\phi) \wedge (-2\sin(2\theta) d\theta \wedge d\psi) = 2\sin(2\theta)d\phi\wedge d\theta\wedge d\psi
	\]
	Since $\alpha\wedge d\alpha$ is non-zero, this agrees with the previous part of the problem.
\end{parts}

\begin{problem}{2}
Suppose $M$ is a smooth manifold and $E\subset TM$ is a distribution. Define:
\[
	\mathcal{I}(E) = \left\{ \omega \in \Omega^\bullet(M) : \omega|_E = 0\right\}.
\]
\end{problem}

\begin{parts}
	\begin{part}{a}
		Prove that $\mathcal{I}(E)\subset \Omega^\bullet_M$ is an ideal.
	\end{part}

	Clearly $\mathcal{I}(E)$ is additively closed. For any $\omega,\eta \in \Omega^\bullet(M)$ with $\omega\in \mathcal{I}(E)$ and $\xi_i\in \Gamma(E)$, recall that
	\[
		(\omega\wedge \eta)(\xi_1,\ldots,\xi_n) = \sum_{\sigma\in \mathfrak{S}_n} \textrm{sgn}(\sigma)\cdot \omega(\xi_{\sigma(1)},\ldots, x_{\sigma(k)})\cdot \eta(x_{\sigma(k+1)}, \ldots, x_{\sigma(n)}),
	\]
	where $|\omega|=k$ and $n=|\omega|+|\eta|$.

	Clearly, every term in this sum will vanish since $\omega|_E=0$ and all vector fields belong to $E$.

	\begin{part}{b}
		Prove that if $E$ has corank $r$ -- that is, if $\dim E_x + r =\dim_x M$ for all $x\in M$ -- then $E$ is locally generated by $r$ independent $1$-forms.
	\end{part}

	It suffices to consider connected manifolds when the dimension is a fixed $n$, since the result is local anyways. Let's work in some chart $U$ with a local frame $\xi_1, \ldots, \xi_{n-r}, \eta_1,\ldots \eta_r$, and $\xi_1,\ldots,\xi_{n-r}$ a local frame for $E$. Let the dual coframe be $\xi^1,\ldots, \xi^{n-r},\ldots, \eta^1,\ldots, \eta^r$, and consider the distribution
	\[
		V = \bigcap_{1\leq i \leq r}\ker \eta^i.
	\]
	At any point $p\in U$, the $n-r$ linearly independent vectors $(\xi_1)_p,\ldots,(\xi_{n-r})_p$ lie in $V_p$ and $(\eta_1)_p,\ldots,(\eta_r)_p$ don't. Since these vectors form a basis for $T_p M$, it follows that $V_p=E_p$ and so the distribution $E$ is generated on $U$ by the independent $1$-forms $\eta^1\,ldots,\eta^r$.

	\begin{part}{c}
		Prove that $\mathcal{I}(E)$ is closed under $d$ if and only if $E$ is integrable.
	\end{part}

	For any $k$-form $\omega\in \Omega^\bullet(M)$ and vector fields $\xi_1,\ldots,\xi_{k+1}\in \Gamma(TM)$, we have
	\[
		\begin{aligned}
			d\omega(\xi_1,\ldots,\xi_{k+1}) & = \sum_{1\leq i \leq k+1} (-1)^{i+1} \xi_i\omega(\xi_1,\ldots,\widehat{\xi_i},\ldots, \xi_{k+1})                                               \\
			                                & \quad+ \sum_{1\leq i<j\leq k+1} (-1)^{i+j+1}\omega([\xi_i, \xi_j], \xi_1, \ldots, \widehat{\xi_i},\ldots, \widehat{\xi_j}, \ldots, \xi_{k+1}),
		\end{aligned}
	\]
	where $\widehat{\xi_i}$ denotes omission of the $i$-th entry. This identity can be proved inductively using the standard commutator relations between $\iota, d,$ and $\mathcal{L}$.

	Now, recall that $E$ is integrable if and only if for any vector fields $\xi_1,\xi_2\in \Gamma(E)$ belonging to $E$, their commutator $[\xi_1, \xi_2]$ belongs to $E$ as well. Supposing $E$ is integrable and that $\omega$ vanishes on $E$, it's clear that $d\omega$ vanishes on $E$ since all the vector fields in the terms of the sum are in $E$, and $\omega$ vanishes on vector fields in $E$.


	Conversely, suppose that $\mathcal{I}(E)$ is closed under $d$. For any forms $\xi,\eta\in \Gamma(E)$, we can locally find 1-forms $\alpha_1,\ldots,\alpha_r$ which locally generate $E$. These forms can be extended to $\Gamma(E^*)$. Then $\alpha_i\in \mathcal{I}(E)$ and hence $d\alpha_i\in \mathcal{I}(E)$ by assumption. However,
	\[
		\alpha_i([\xi, \eta]) = \eta \alpha_i(\xi) - \xi\alpha_i(\eta) - d\alpha_i(\xi,\eta).
	\]
	All the terms on the right hand side must vanish by assumption, and so $\alpha_i([\xi,\eta])=0$ for all $i$. From this it's immediately implied that $[\xi,\eta]\in \Gamma(E)$ locally. We can do this across the entire manifold so $E$ is integrable.

	\begin{part}{d}
		Consider the distribution $E$ on $\mathbb{A}^3_{x,y,z}$ spanned by the vector fields $\partial/\partial x$ and $x\partial/\partial y + \partial/\partial z$. Show that $E$ is not integrable. Show that any point $(x,y,z)\in \mathbb{A}$ may be joined to the origin by a piecewise smooth curve whose tangent line belongs to $E$.
	\end{part}

	Letting $\xi=\pps{x}$ and $\eta = x\pps{y} + \pps{z}$, we can compute the commutator $[\xi, \eta]$ as
	\[
		[\xi,\zeta] = \left[\pp{}{x}, x\pp{}{y} + \pp{}{z}\right] = \pp{}{y}.
	\]
	This proves that $E$ cannot integrable since this commutator can clearly not be expressed as a linear combination of $\xi$ and $\eta$.

	Next, let's show that any point in $\mathbb{A}^3$ can be joined to the origin by a piecewise smooth curve whose tangent lines form a subbundle of $E$. The flows associated to the vector fields $\xi$ and $\eta$ are
	\[
		\varphi_t(x,y,z)= ( x+t, y, z)\qtq{and}\psi_t(x,y,z) = (x,xt+y, t+z).
	\]
	Thus, the piecewise smooth integral curve is given by travelling along $\xi$ for time $y/z-x$ and then travelling along $\eta$ for time $-y/z$. If $z=0$, we can follow $\xi$ for time $-x$ and then $\psi$ for time $1$. Then, we repeat the procedure for the nonzero $z$ case.
\end{parts}

\begin{problem}{3}
Example or proof of nonexistence: A codimension 1 foliation on the sphere $S^4$.
\end{problem}

\begin{solution}
	Suppose for the sake of contradiction that exists a foliation $\mathcal{F}$ of codimension $1$ on $S^4$. Let $E_\mathcal{F}$ be the associated distribution of codimension $1$. Using the standard metric on $S^4$, the orthogonal complement $E_\mathcal{F}^\perp$ gives us a line field on $S^4$. However, $\chi(S^4)=2$, so it cannot admit a line field. This is a contradiction, so a codimension 1 foliation does not exist.
\end{solution}

\begin{problem}{4}
The Frobenius tensor.
\end{problem}

\begin{parts}
	\begin{part}{a}
		Let $P,Q: \mathbb{A}^2 \to \R$ be smooth functions. Define the $2$-dimensional distribution $E$ on $\mathbb{A}^2_{x,y}\times \R_z$ with
		\[
			E_{(x,y,z)} = \textrm{span}\left\{\frac{\partial}{\partial x} + P \frac{\partial}{\partial z}, \frac{\partial}{\partial y} + Q \frac{\partial}{\partial z}\right\}.
		\]
		Compute the Frobenius tensor of $E$.
	\end{part}

	Since $E$ is spanned by the vector fields
	\[
		\xi_1 = \pp{}{x} + P\pp{}{z},\qtq{and} \xi_2 = \pp{}{y}+Q\pp{}{z}
	\]
	it suffices to compute their commutator in order to determine the Frobenius tensor. This computation yields
	\[
		\begin{aligned}
			\phi_E(\xi_1, \xi_2)=
			 & = \left[\pp{}{x} + P\pp{}{z}, \pp{}{y}+Q\pp{}{z} \right]\mod E \\
			 & = \left(\pp{Q}{x} - \pp{P}{y}\right)\pp{}{z} \mod E.
		\end{aligned}
	\]
	Thus, if vector fields $\eta_1, \eta_2$ in $E$ have coefficients $\eta_1 = \alpha \xi_1 + \beta \xi_2$ and $\eta_2 = \gamma \xi_1 + \kappa \xi_2$, then
	\[
		\phi_E(\eta_1, \eta_2) = (\alpha\kappa - \beta\gamma)\left(\pp{Q}{x} - \pp{P}{y}\right)\pp{}{z} \mod E.
	\]

	\begin{part}{b}
		Suppose $X$ is a manifold and $G$ a Lie group. Let $\{\theta^i\}$ be a basis of left-invariant $1$-forms on $G$ and suppose
		\[
			d\theta^i + \frac{1}{2}c^i_{jk} \theta^j\wedge \theta^k = 0
		\]
		for constants $c^i_{jk}$. Let $\{\theta^i_X\}$ be $1$-forms on $X$. Consider the ideal of differential forms on $X\times G$ generated by $\pi_G^*\theta^i - \pi^*_X \theta^i_X$, where $\pi_X : X \times G \to X$ and $\pi_G : X\times G \to G$ are projections. Prove that this ideal is closed under $d$ if and only if
		\[
			d\theta^i_X + \frac{1}{2}c^i_{jk} \theta^j_X \wedge \theta^k_X = 0.
		\]
	\end{part}

	Let $\mathcal{I}$ denote the ideal. To simplify notation, let $x^i = \theta^i_X$ and $X^i=\pi_X^* \theta^i_X$. Similarly, let $g^i = \theta^i$ and $G^i = \pi_G^*\theta^i$.
	Writing the given relation in this form, we can obtain expressions
	\[
		dg^i + \frac{1}{2}c^i_{j k} g^j\wedge g^k = 0\quad\implies\quad dG^i + \frac{1}{2}c^i_{j k} G^j\wedge G^k,
	\]
	by the naturality of $d$ and $\wedge$.
	Expanding the differential of a generator, we get
	\[
		\begin{aligned}
			d(G^i - X^i)
			 & = dG^i - dX^i                                 \\
			 & = -\frac{1}{2}c^i_{j k} G^j \wedge G^k - dX^i \\
		\end{aligned}
	\]
	Now assume that the have the identity $dx^i + (c^i_{j k} /2)\,x^j \wedge x^k = 0$. Then by naturality of $d$ and $\wedge$ we have the identity $dX^i + (c^i_{j k} / 2)\,X^j \wedge X^k=0$. Then expanding the differential of a generator, we have
	\[
		\begin{aligned}
			d(G^i - X^i)
			 & = dG^i - dX^i                                                                  \\
			 & = -\frac{1}{2}c^i_{j k}G^j \wedge G^k + \frac{1}{2}c^i_{j k} X^j \wedge X^k    \\
			 & = -\frac{1}{2}c^i_{j k}\left( G^j\wedge G^k + X^j \wedge X^k\right)            \\
			 & = -\frac{1}{2}c^i_{j k}\left( (G^j-X^j)\wedge G^k - (G^k-X^k)\wedge G^j\right)
		\end{aligned}
	\]
	This is a sum of elements of the ideal, so the ideal is closed under $d$.

	Next, let's suppose that the ideal is closed under $d$. We know that:
	\[
		\begin{aligned}
			 & d(G^i-X^i) = dX^i + \frac{1}{2}c^i_{j k} G^j\wedge G^k\in \mathcal{I}      & \quad\quad \textrm{subtracting }\frac{1}{2}c^i_{j k} (G^j - X^j)\wedge G^k - \frac{1}{2}c^i_{j k} X^j\wedge (G^k - X^k) \\
			 & \quad\implies dX^i + \frac{1}{2} c^i_{j k} X^j \wedge X^k \in \mathcal{I}.
		\end{aligned}
	\]
	We would like to show that $dx^i+(c^i_{j k} / 2) x^j\wedge x^k=0$, and so far, we've shown that its pullback under $\pi_X$ is in $\mathcal{I}$. Let's let $\omega = dx^i+(c^i_{j k} / 2) x^j\wedge x^k$. Using the generators of $\mathcal{I}$, we can find some $1$-forms $\alpha^q\in \Omega^{1}(X\times G)$ such that:
	\[
		\pi_X^*\omega = \sum_q(G^q - X^q)\wedge \alpha^q.
	\]
	Let's write $\alpha^q = a_{k,q}\,G^k + b_{k,q}\,\beta^{k,q}$ for some forms $\beta^{k,q}\in \Omega^{1,0}(X\times G)$ and $0$-forms $a_{k,q},b_{k,q}$ -- we can do this because $g^i$ form a basis for $\Omega^1(G)$ and hence $G^i$ form a basis for $\Omega^{0,1}(X\times G)$. Then we have
	\[
		\begin{aligned}
			\pi_X^*\omega = dX^i + \frac{1}{2}c^i_{j k} X^j\wedge X^k
			 & = \sum_{q,k} (G^q - X^q)\wedge \alpha^q                                                                                          \\
			 & = \sum_{q,k} (G^q - X^q)\wedge a_{k,q} G^q  + (G^q-X^q)\wedge b_{k,q} \beta^{k,q}                                                \\
			 & = \sum_{q,k} G^q\wedge a_{k,q} G^q + (G^q\wedge b_{k,q} \beta^{k,q} - X^q\wedge a_{k,q} G^{k,q}) - X^q\wedge b_{k,q} \beta^{k,q}
		\end{aligned}
	\]
	Since the original form $\pi^*_X\omega$ is in $\Omega^{2,0}(X\times G) \subset \Omega^2(X\times G)$, the terms involving $G^q$ must vanish, so in particular $a_{k,q}=0$ and consequently $b_{k,q}=0$. However, this means that $\pi_X^*\omega=0$, which also means that $\omega=0$. Thus,
	\[
		\omega = d\theta_X^i + \frac{1}{2}c^i_{j k} \theta^j_X \wedge \theta^k_X = 0.
	\]


	\begin{part}{c}
		Compute the Frobenius tensor of the distribution in (b) defined as the simultaneous kernel of the $1$-forms $\pi_G^*\theta^i - \pi_X^*\theta^i_X$.
	\end{part}

	\textcolor{red}{
		Let $E$ be the distribution on $X\times G$ defined as the intersection
		\[
			E = \mathop{\bigcap}_i \ker (G^i - X^i).
		\]
		The Frobenius tensor is defined as the bilinear map
		\[
			\definefunction{\phi_E}{\Gamma(E)\times \Gamma(E)}{\Gamma(T(X\times G)/E)}{(\xi, \eta)}{[\xi,\eta]\mod E.}
		\]
		Suppose $\xi$ is a vector field in $E$, which equivalently means that it is a vector field on $X\times G$ with $g^i(\xi_G)=x^i(\xi_X)$, where $\xi_G$ and $\xi_X$ are the projections of $\xi$ to $G$ and $X$ respectively. For any two vector fields $\xi$ and $\eta$ in $E$, we have relations
		\[
			\begin{aligned}
				dg^i(\xi_G,\eta_G) + \frac{1}{2}c^i_{j k} g^j(\xi_G)\wedge g^k(\eta_G)
				                                                                                                                 & = 0 \\
				\xi_G g^i(\eta_G) - \eta_G g^i(\xi_G) - g^i([\xi_G, \eta_G]) + \frac{1}{2}c^i_{j k} g^j(\xi_G)\wedge g^k(\eta_G) & = 0
			\end{aligned}
		\]
		Unsure how to finish this...
	}
\end{parts}

% \begin{problem}{5}
%   This is a collection of exercises on the Maurer-Cartan form.
% \end{problem}
%
% \begin{parts}
%   \begin{part}{a}
%     Let $G$ be a Lie group with Maurer-Cartan form $\theta$. Compute $R_g^*\theta$ for $g \in G$. Do this first for a matrix group, where you can write $\theta = g^{-1}dg$ for $g : G \to M_n(\R)$ the natural matrix-valued function on a matrix group.
%   \end{part}
%
%   \begin{part}{b}
%     Let $G$ be a Lie group and suppose $T$ is a \emph{right} $G$-torsor. Show that the Maurer-Cartan form on $G$ transports to a canonical element of $\Omega^1(T; \mathfrak{g})$. Can you give a prose definition of this Maurer-Cartan $1$-form on the torsor? What is the Maurer-Cartan equation? What is the pullback of the Maurer-Cartan $1$-form by an element of $G$ acting on $T$?
%   \end{part}
%
%   \begin{part}{c}
%     Let $V$ be an $n$-dimensional real vector space and $\scB(V)$ the right $\GL_n$-torsor of bases. Let $\Theta^i_j$ be the Maurer-Cartan forms in the standard basis of the Lie algebra of $\GL_n$. Suppose $b(t)$ is a smooth curve in $\scB(V)$. Write the basis $b(t)$ as $\{e_1(t),\ldots, e_n(t)\}$ and the dual basis as $\{e^1(t),\ldots, e^n(t)\}$. Prove that
%     \[
%       \Theta^i_j(\dot{b}) = \langle e^i(0), \dot{e}_j(0)\rangle.
%     \]
%     Heuristically then, $\Theta^i_j$ measures the instantaneous motion of $e_j$ in the direction of $e_i$, where `direction' is determined by the entire basis $e_1,\ldots, e_n$. This interpretation is important!
%   \end{part}
%
%   \begin{part}{d}
%     Let $A$ be an $n$-dimensional real affine space and $\scB(A)$ the right $\Aff_n$-torsor of bases of the underlying vector space at all points of $A$. Let $\theta^i$, $\Theta^i_j$ be the Maurer-Cartan forms in the standard basis of the Lie algebra of $\Aff_n$. (Define this basis: the single index is for infinitesimal translations, and the double index for infinitesimal linear transformations, as in (c).) Suppose $b(t)$ is a smooth curve in $\scB(A)$ which projects to the curve $x(t)$ in $A$, and write the underlying basis of $V$ as in (c). Prove that
%     \[
%
%     \]
%   \end{part}
% \end{parts}

\end{document}
