\documentclass{../../templates/lkx_pset}

\usepackage[T1]{fontenc}
\RequirePackage{mlmodern}

\title{Math 230a Problem Set 11}
\author{Lev Kruglyak}
\due{December 4, 2024}

\collaborator{AJ LaMotta}
\collaborator{Ignasi Vicente}

\input{../math230a.sty}

\begin{document}
\maketitle

% \begin{problem}{1}
% Let $X$ be a Riemannian manifold and suppose $\{x_1,\ldots, x_N\}\subset X$ is a finite subset. In lecture, I constructed a manifold with boundary $\widehat{X}$ and map $p : \widehat{X}\to X$ that is a diffeomorphism on the complement of $\bigsqcup_i p^{-1}(x_i)$. Furthermore, if $\xi$ is a vector field on $X$ with a transverse zero at each $x_i$ and no other zero, then the normalization $s = \xi/\|\xi\| : X\setminus\{x_1,\ldots,x_N\} \to S(TX)$ extends over $\widehat{X}$.
% Explore this construction in detail when $\dim X=1$.
% \end{problem}
%
\begin{problem}{2}
Consider Chern-Simons-Weil forms for $G=\U_1$ the circle group of unit norm complex numbers. The Lie algebra is $\mathfrak{g}=i\R$.
\end{problem}
\begin{parts}
	\begin{part}{(a)}
		For each $k\in \Z^{>0}$ identify the vector space of degree $k$ symmetric $G$-polynomials on $\mathfrak{g}$.
	\end{part}

	Since $\U_1$ is abelian, the adjoint representation is trivial.
	Now as $\mathfrak{u}_1=i\R$, the vector space of degree $k$ symmetric $G$-invariant polynomials is just the space of symmetric polynomials
	\[
		\R[-ix_1, \ldots, -ix_k]^{\mathfrak{S}_k}.
	\]

	\begin{part}{(b)}
		If $\pi : P \to X$ is a principal $G$-bundle and $h$ is a $G$-invariant polynomial of degree $k$, then the Chern-Simons form restricts to a closed $(2k-1)$-form on each fiber of $\pi$, and this form may be identified with a bi-invariant form on $G$. Identify this form for $k=1$. For which $h$ is the integral of this form an integer?
	\end{part}

	Recall from the notes that the pullback of the Chern-Simons form $\alpha$ to a fiber is given by
	\[
		\iota_x^*\alpha = c_k\cdot h(\theta_{P_x}\wedge [\theta_{P_x}\wedge \theta_{P_x}]^{\wedge(k-1)})\quad\textrm{where}\quad c_k = k\int_0^1 dt\left(\frac{t^2-t}{2}\right)^{k-1}
	\]
	and $\theta_{P_x}$ is the Maurer-Cartan form. In out case of $k=1$, we have $c_k=1$ and $h(x)=-i\kappa\cdot x$ so the restricted Chern-Simons form is $\iota_x^*\alpha = -i\kappa\theta_{P_x}.$ This can be identified with the bi-invariant form $-i\kappa\cdot \theta_{U_1} = \kappa\,d\theta\in \Omega^1(\U_1)$. The integral of this form is
	\[
		\int_{P_x} \iota_x^*\alpha= \int_{\U_1} -i\kappa\cdot \theta_{\U_1} = \int_0^{2\pi} \kappa \cdot d\theta = 2\pi \kappa.
	\]
	For this to be an integer, the polynomial $h$ must take the form
	\[
		h(x) = \frac{n}{2\pi i}\cdot x\quad\textrm{for}\quad n\in \Z.
	\]

	\begin{part}{(c)}
		Consider the Hopf bundle $\pi : S^3 \to S^2$, which is a principal $G$-bundle. Construct a connection. Compute the integral of the Chern-Weil form over $S^2$. For which $h$ is this an integer?
	\end{part}

	Recall that $\SU_2$ consists of unitary complex $2\times 2$ matrices with determinant $1$, and the Lie algebra $\su_2$ consists of $2\times 2$ skew-Hermitian traceless matrices. We can explicitly parametrize them by
	\[
		\SU_2 = \left\{
		\begin{pmatrix} \alpha & -\overline{\beta}\\ \beta & \overline{\alpha}\end{pmatrix}
		:
		\begin{array}{l}\alpha,\beta\in \C\\|\alpha|^2+|\beta|^2=1\end{array}
		\right\}
		\qtq{and}
		\su_2 = \left\{
		\begin{pmatrix}
			-i\theta & \overline{z} \\
			-z       & i\theta
		\end{pmatrix}
		:
		\theta\in \R, z\in \C
		\right\}.
	\]
	The Lie algebra $\su_2$ can be given a basis of Pauli matrices
	\[
		\sigma_1 = \frac{1}{\sqrt{2}}\begin{pmatrix}0 & 1\\ -1 & 0\end{pmatrix},\quad
		\sigma_2 = \frac{1}{\sqrt{2}}\begin{pmatrix}0 & -i\\ -i & 0\end{pmatrix},\qtq{and}
		\sigma_3 = \frac{1}{\sqrt{2}}\begin{pmatrix}-i & 0\\ 0 & i\end{pmatrix}.
	\]
	Note that the commutators are $[\sigma_1, \sigma_2] = -\sigma_3$, $[\sigma_2, \sigma_3]=-\sigma_1$, and $[\sigma_3, \sigma_1]=-\sigma_2$. We include the normalization factor of $1/\sqrt{2}$ so that we have
	\[
    \textrm{Tr}(\sigma_1^2) = \Tr(\sigma_2^2) = \Tr(\sigma_3^2) = 1
	\]
	or in other words, $\sigma_1,\sigma_2,\sigma_3$ is an orthonormal basis for $\su_2$ under the inner product $\langle X, Y\rangle=\Tr(X^\dagger Y)$.

	% Exponentiating these generators gives us 
	% \[
	%   e^{t\sigma_1} = \begin{pmatrix}\cos(t) & -\sin(t)\\ \sin(t) & \cos(t)\end{pmatrix},\quad
	%   e^{t\sigma_2} = \begin{pmatrix}\cos(t) & i\sin(t)\\ i\sin(t) & \cos(t)\end{pmatrix},\qtq{and}
	%   e^{t\sigma_3} = \begin{pmatrix}e^{it}&0\\ 0 & e^{-it}\end{pmatrix}.
	% \]

	Now let's construct a connection $\Theta\in \Omega^1(\SU_2; \mathfrak{u}_1)$ on $\pi : S^3 \to S^2$. Let's start with the Maurer-Cartan form $\theta_{\SU_2}\in\Omega^1(\SU_2; \su_2)$. The Maurer-Cartan form can be written as $\theta_{\SU_2} = \theta^1\sigma_1 + \theta^2\sigma_2+\theta^3\sigma_3$ for some forms $\theta^i\in \Omega^1(\SU_2)$. These forms must satisfy the Maurer-Cartan equations:
	\[
		d\theta^i + \frac{1}{2}\sum_{j,k} c^i_{j k} \theta^j\wedge \theta^k = 0 \quad\implies\quad \begin{array}{l}
			d\theta^1 = \theta^2\wedge \theta^3 \\
			d\theta^2 = \theta^3\wedge \theta^1 \\
			d\theta^3 = \theta^1\wedge \theta^2 \\
		\end{array}
	\]
	Let's split $\su_2 = \mathfrak{m}\oplus \mathfrak{u}_1$ by letting $\mathfrak{m}$ be the span of $\sigma_1$ and $\sigma_2$, and with $\mathfrak{u}_1$ the span of $\sigma_3$. This is a reductive structure on the symmetric space $\SU_2/\U_1$. Let's now let $\Theta$ be the projection of $\theta_{\SU_2}$ onto $\mathfrak{u}_1$. We can thus write $\Theta = \theta^3\sigma_3 = i\cdot \theta^3$. Since $\U_1$ is abelian, the curvature is
	\[
		\Omega = d\Theta + \frac{1}{2}[\Theta\wedge \Theta] = d\Theta = i(d\theta^3) = i\cdot (\theta^1\wedge \theta^2) \in \Omega^2(\SU_2; \mathfrak{u}_1).
	\]
	Note that all Chern-Weil forms vanish for $k>1$, since $\Omega\wedge \Omega=0$, as it is a $4$-form on the $3$-manifold $\SU_2$. When $k=1$, a polynomial $h\in (\textrm{Sym}^k \mathfrak{u}_1^*)^{\U_1}$ can be written as $h(x)=-i\kappa\cdot x$ for some $c_k\in \R$. In this case, the Chern-Weil form is
	\[
		\omega = h(\Omega) = \kappa\cdot(\theta^1\wedge \theta^2).
	\]
	This is a closed $2$-form on $\SU_2$ since $\omega = d(\kappa\cdot \theta^3)$. It also must descend to some $2$-form $\widetilde{\omega}\in \Omega^2(S^2)$, i.e. we have $\omega=\pi^*\widetilde{\omega}$. We would like to calculate the integral of $\widetilde{\omega}$ over $S^2$. It's clear that $\widetilde{\omega}$ is an area form on $S^2$, so all that remains is to compute the scale factor with respect to the standard area form on the unit sphere.

	Let's first write the Hopf fibration as
	\[
		\pi\begin{pmatrix}\alpha & -\overline{\beta}\\ \beta & \overline{\alpha}\end{pmatrix} = \begin{pmatrix}2\cdot \Re(\alpha \overline{\beta}) \\ 2\cdot \Im(\alpha\overline{\beta}) \\ |\alpha|^2 - |\beta|^2\end{pmatrix}\in S^2\subset \R^3
	\]
	Writing $\alpha = x_1+iy_1$ and $\beta = x_2+iy_2$, the differential of this map now takes the form
	\[
		d\pi_{(\alpha,\beta)} = \begin{pmatrix}2x_2 & 2y_2 & 2x_1 & 2y_1 \\ -2y_2 & 2x_2 & 2y_1 & -2x_1 \\ 2x_1 & 2y_1 & -2x_2 & -2y_1\end{pmatrix}.
	\]
	The differential of the basis for $T_e\SU_2$ give by the Pauli matrices thus becomes:
	\[
		d\pi_{(1,0)} \sigma_1 = \begin{pmatrix}-\sqrt{2}\\0\\0\end{pmatrix},\quad
		d\pi_{(1,0)} \sigma_2 = \begin{pmatrix}0\\\sqrt{2}\\0\end{pmatrix},\quad
		d\pi_{(1,0)} \sigma_3 = \begin{pmatrix}0\\0\\0\end{pmatrix}.
	\]
	Since these tangent vectors have length $\sqrt{2}$, it follows that $\theta^1\wedge \theta^2$ descends to $(1/\sqrt{2})(1/\sqrt{2})dA$ where $dA$ is the area form on the unit sphere $S^2$. In particular, this implies that
	\[
		\int_{S^2} \widetilde{\omega} = \frac{\kappa}{2}\int_{S^2}dA = \kappa\cdot 2\pi.
	\]
	Thus, the only polynomials for which the integral is an integer are
	\[
		h(x) = \frac{n}{2\pi i}\cdot x\quad \textrm{for} \quad n\in \Z.
	\]

	\begin{part}{(d)}
		Now consider $k=2$, so the Chern-Weil form has degree $4$. Construct a nontrivial principal $G$-bundle over $S^2\times S^2$ by first taking the Cartesian product of the Hopf bundle with itself to form a principal $(G\times G)$-bundle over $S^2\times S^2$, then use the homomorphism $G\times G\to G$ to form the associated principal $G$-bundle. Compute the integral of the Chern-Weil form. For which $h$ is the answer an integer?
	\end{part}

	Let's begin with the principal $(\U_1\times \U_1)$-bundle $\pi \times \pi : \SU_2\times_{S^2\times S^2} \SU_2 \to S^2\times S^2$ -- for brevity let's call the total space $E$. Let $\pi_1,\pi_2 : E \to \SU^2$ be respective projection maps, and let $\widetilde{\pi_i}=\pi\circ \pi_i$ be the composition with the Hopf fibration. The induced connection on this bundle can be written as
	\[
		\Theta = (i\cdot \pi_1^*\theta^3) \oplus (i\cdot \pi_2^*\theta^3) \in \Omega^1(E; \mathfrak{u}_1\oplus \mathfrak{u}_1)
	\]
	By the results of the previous problem, the curvature of this connection is then
	\[
		\Omega = (i\cdot \widetilde{\pi}_1^*\, dA) \oplus (i \cdot \widetilde{\pi}_2^*\, dA) \in \Omega^2(E; \mathfrak{u}_1\oplus\mathfrak{u}_1),
	\]
	where $dA$ is the area form on $S^2$.


	Now we want to understand the pushforward of $\Theta$ to the principal $\U_1$-bundle $E'=E\times_{\U_1\times \U_1} \U_1$ associated to $E$ under the multiplication homomorphism $\mu : \U_1\times \U_1\to \U_1$. The differential $\mu_* : \mathfrak{u}_1\oplus \mathfrak{u}_1 \to \mathfrak{u}_1$ is the addition map. Let $\overline{\theta}_i^j\in \Omega^1(E')$ be the pushforwards of $\theta^i_j\in \Omega^1(E)$ to the associated bundle. Then, the connection and curvature of the associated bundle is
	\[
		\Theta' = i\cdot (\overline{\theta}_1^3 + \overline{\theta}_2^3)\qtq{and}\Omega' = 2i\cdot (
		\overline{\theta}^1_1\wedge \overline{\theta}^2_1+
		\overline{\theta}^1_2\wedge \overline{\theta}^2_2
		).
	\]
	A polynomial $h\in (\textrm{Sym}^2\mathfrak{u}_1^*)^{\U_1}$ can be written as $h(x) = -\kappa \cdot xy$ so the Chern-Weil form is
	\[
		\begin{aligned}
			\omega = h(\Omega'\wedge \Omega') = -\kappa \cdot \Omega'\wedge \Omega'
			 & = 4\kappa\cdot
			(\overline{\theta}^1_1\wedge \overline{\theta}^2_1+ \overline{\theta}^1_2\wedge \overline{\theta}^2_2)
			(\overline{\theta}^1_1\wedge \overline{\theta}^2_1+ \overline{\theta}^1_2\wedge \overline{\theta}^2_2) \\
			 & = 8\kappa\cdot
			\overline{\theta}^1_1\wedge \overline{\theta}^2_1\wedge
			\overline{\theta}^1_2\wedge \overline{\theta}^2_2.
		\end{aligned}
	\]
	Recall that $\overline{\theta}^1_i\wedge \overline{\theta}^2_i$ descends to a form $\widetilde{\omega}_i$ on the $i$-th $S^2$ term in $S^2\times S^2$ with total integral $2\pi$. Since $\omega$ descends to $8\kappa\cdot \widetilde{\omega}_1 \wedge \widetilde{\omega}_2$ on $S^2\times S^2$, we can compute
	\[
		\int_{S^2\times S^2} 8\kappa\cdot \widetilde{\omega}_1\wedge \widetilde{\omega}_2 = 8\kappa\cdot\left(\int_S^2 \widetilde{\omega}_1\right)^2 = 32\pi^2\kappa.
	\]
	Thus, the only $h$ for which the integral is an integer are
	\[
		h(x) = -\frac{n}{32\pi^2}\cdot x^2\qtq{for}n\in \Z.
	\]

	\begin{part}{(e)}
		Continuing with $k=2$, take the base $4$-manifold to be $\CP^2$. For which $h$ do you find an integer when you integrate the Chern-Weil form?
	\end{part}

	In this case, the canonical bundle to consider is the generalized Hopf fibration
	\[
		\U_1 \to S^5 \to \CP^2
	\]
	where the $5$-sphere $S^5$ is considered as the space of triples of complex numbers in $\C^3$ with norm $1$. The map sends such a triple $(\alpha, \beta, \gamma)$ to the same homogeneous coordinates $[\alpha:\beta:\gamma]$.

	\todo{unsure how to proceed}
\end{parts}

\begin{problem}{3}
Now consider $G=\SU_2$.
\end{problem}
\begin{parts}
	\begin{part}{(a)}
		Identify the space of $G$-invariant polynomials of degree $2$ on $\mathfrak{g}$.
	\end{part}

	Let's use the same bases from the previous problem. First, note that the adjoint representation of $\SU_2$ on $\su_2$ factors through the double cover $\SU_2 \to \SO_3$.
	This immediately implies that there are no nonzero $\SU_2$-invariant polynomials of degree $k=1$. In the next case of $k=2$, this is the space of symmetric bilinear forms on $\su_2$ which are invariant under the action of $\SO_3$.
	Up to scaling, the only such bilinear form is $\langle X, Y\rangle = \Tr(X^\dagger Y)$. Thus, we can write any $h\in (\Sym^2\mathfrak{su}_2^*)^{\SU_2}$ as $h(X\wedge Y) = \kappa\cdot \Tr(X^\dagger Y)$ for some $\kappa\in \R$. (Technically this Hermitian transpose is unnecessary since all elements in $\mathfrak{su}_2$ are skew-Hermitian and so this is a scalar multiple of the bilinear form $\Tr(XY)$)

	\begin{part}{(b)}
		Write an explicit formula for the Chern-Simons $3$-form of a connection. Use matrix multiplication in your formula rather than the Lie bracket.
	\end{part}

	If $h\in (\Sym^2\mathfrak{su}_2^*)^{\SU_2}$, then the Chern-Simons $3$-form of a connection $\Theta\in \Omega^1(P; \su_2)$ is:

	% \[
 %    \textrm{CS}_3(\Theta) = 2\cdot h(\Theta\wedge \Omega) = 2\kappa\cdot \Tr\left(\Theta\, d\Theta + \frac{1}{2}\Theta[\Theta\wedge\Theta]\right) = 2\kappa\left(\Tr(\Theta\, d\Theta) + \frac{1}{2}\Tr(\Theta[\Theta\wedge\Theta])\right)
	% \]
	\[
		\textrm{CS}_3(\Theta) = 2\int_0^1 dt\,h(\Theta \wedge \Omega_t)\quad\textrm{where}\quad \Omega_t = t\,d\Theta + \frac{t^2}{2}[\Theta\wedge \Theta].
	\]
	Since $h$ is a symmetric bilinear form, we get
	\[
		\begin{aligned}
		\textrm{CS}_3(\Theta) = 2\int_0^1 t\cdot h(\Theta\wedge d\Theta) + \frac{t^2}{2}\cdot h\left(\Theta\wedge[\Theta\wedge\Theta]\right)\,dt 
		&= h(\Theta\wedge d\Theta) + \frac{1}{3}h(\Theta\wedge [\Theta\wedge \Theta])
		\end{aligned}
	\]
	Let's suppose $h(X\wedge Y)=\Tr(X^\dagger Y)$, since all other forms under consideration are scalar multiples of this bilinear form. Since $\su_2$ is skew-hermitian, it follows that $h(\Theta\wedge d\Theta) = \Tr(\Theta^\dagger d\Theta)= - \Tr(\Theta d\Theta)$. For the term involving the Lie bracket, we write $\Theta = \theta^i\sigma_i$ where $\sigma_i$ are the Pauli matrices defined in the previous problem. Then by applying the commutator rules we get
	\[
		h(\Theta\wedge [\Theta\wedge \Theta]) = \theta^k\wedge \theta^i\wedge \theta^j\Tr(\sigma_k^\dagger [\sigma_i,\sigma_j]) = 6\cdot \theta^1\wedge \theta^2\wedge \theta^3.
	\]
	Similarly, expanding the trace of a power of the connection, we get
	\[
		\Tr(\Theta^3) = \theta^i\wedge\theta^j\wedge \theta^k\Tr(\sigma_i\sigma_j\sigma_k) = \theta^1\wedge \theta^2\wedge \theta^3.
	\]
	Combining everything together, we get the Chern-Simons $3$-form of an $\SU_2$-connection solely in terms of matrix multiplication:
	\[
		\textrm{CS}_3(\Theta) = \Tr(\Theta^\dagger d\Theta + \Theta^\dagger[\Theta\wedge\Theta]/3) =
		\Tr(-\Theta d\Theta) + 2\Tr(\Theta^3)=
		\Tr(2\Theta^3-\Theta d\Theta).
	\]

	% \begin{part}{(c)}
	% 	Use the homomorphism $\U_1 \to \SU_2$ to construct $\SU_2$-connections from $\U_1$-connections. How are the Chern-Simons forms related?
	% \end{part}

	% Since the homomorphism $\iota : \U_1\to \SU_2$ induces the Lie algebra homomorphism $\mathfrak{u}_1 \to \su_2$ sending $i$ to $\sigma_3$, it follows that any $\U_1$-connection $\Theta\in \Omega^1(P; \mathfrak{u}_1)$ is associated to a connection $\iota(\Theta)\in \Omega^1(\iota(P); \su_2)$ such that $\iota(\Theta)= \theta\sigma_3$.

	% \begin{part}{(d)}
	% 	Investigate integrality of the integral of Chern-Weil forms for $k=2$.
	% \end{part}
\end{parts}

\begin{problem}{4}
  Let $\Sigma\subset\mathbb{E}^3$ be a closed surface that is a submanifold of Euclidean $3$-space. Prove that $\Sigma$ has a point of positive Gauss curvature.
\end{problem}
\begin{solution}
  % Suppose for the sake of contradiction that $\Sigma$ has non-positive Gauss curvature everywhere. By the Gauss-Bonnet theorem, this means that $\chi(\Sigma)\leq 0$. If $\chi(\Sigma)=0$, this would mean that $\Sigma$ is a surface of constant curvature zero -- impossible since this would mean that it was a plane which cannot be a closed manifold. Since $\Sigma$ is embeddable in $\mathbb{E}^3$, it must be orientable. These constraints imply that $\Sigma$ must topologically be a genus $g$-surface for $g\geq 2$.
  Since $\Sigma$ is compact, it must be bounded by a sphere in $\mathbb{E}^3$. Without loss of generality, we can shrink the sphere so that it intersects $\Sigma$ tangentially at some point $p\in \Sigma$. Note that a non-tangential intersection point of this type is impossible, since the sphere would no longer bound the surface. Then the Gauss curvature at this point $K_p$ must be positive, since otherwise following a geodesic path would allow one to reach points of $\Sigma$ outside of the sphere -- a contradiction.
\end{solution}

\begin{problem}{5}
Let $G$ be a Lie group, let $H\subset G$ be a closed Lie subgroup, and assume that the homogeneous manifold $G/H$ has a reductive structure, i.e. an $H$-invariant splitting $\mathfrak{g}=\mathfrak{p}\oplus\mathfrak{h}$ as vector spaces.
\end{problem}

\begin{parts}
	\begin{part}{(a)}
		Recall the definition of the canonical connection on the principal $H$-bundle $\pi : G\to G/H$.
	\end{part}

	A connection on $\pi$ is a form $\Theta \in \Omega^1(G; \mathfrak{h})$. The canonical choice is to start with the Maurer-Cartan form $\theta_G\in \Omega^1(G; \mathfrak{g})$ and set $\Theta =\pi_\mathfrak{h} \theta_G$ where $\pi_\mathfrak{h} : \mathfrak{g} \to \mathfrak{h}$ is the projection with kernel $\mathfrak{p}$.

	\begin{part}{(b)}
		Compute its curvature.
	\end{part}
	The Maurer-Cartan equation states that
	\[
    d\theta_G + \frac{1}{2}[\theta_G\wedge\theta_G]=0.
	\]
  Combining this with the curvature of the connection and the facts that $[\mathfrak{h},\mathfrak{h}]\subset \mathfrak{h}$ and $[\mathfrak{h},\mathfrak{p}]\subset\mathfrak{p}$, we get:
  \[
    \begin{aligned}
      \Omega 
      &= d\Theta + \frac{1}{2}[\Theta\wedge\Theta]\\
      &= d\pi_\mathfrak{h}\theta_G + \frac{1}{2}[\pi_\mathfrak{h}\theta_G\wedge \pi_\mathfrak{h}\theta_G]\\
      &= \pi_\mathfrak{h}d\theta_G + \frac{1}{2}[\pi_\mathfrak{h}\theta_G\wedge \pi_\mathfrak{h}\theta_G]\\
      &= -\frac{1}{2}\pi_\mathfrak{h}[\theta_G\wedge\theta_G] + \frac{1}{2}[\pi_\mathfrak{h}\theta_G\wedge \pi_\mathfrak{h}\theta_G]\\
      &= -\frac{1}{2}\pi_\mathfrak{h}[
      (\pi_\mathfrak{h}\theta_G + \pi_\mathfrak{p}\theta_G)\wedge(\pi_\mathfrak{h}\theta_G + \pi_\mathfrak{p}\theta_G)
    ] + \frac{1}{2}[\pi_\mathfrak{h}\theta_G\wedge \pi_\mathfrak{h}\theta_G]\\
      &= -\frac{1}{2}\pi_{\mathfrak{h}}[\pi_{\mathfrak{h}} \theta_G \wedge \pi_{\mathfrak{h}}\theta_G]
         -\frac{1}{2}\pi_{\mathfrak{h}}[\pi_{\mathfrak{p}} \theta_G \wedge \pi_{\mathfrak{p}}\theta_G]
         +\frac{1}{2}[\pi_{\mathfrak{h}} \theta_G \wedge \pi_{\mathfrak{h}}\theta_G]\quad\quad(\textrm{since } \pi_\mathfrak{h} [\pi_\mathfrak{h}\theta_G\wedge \pi_\mathfrak{p}\theta_G]=0)\\
      &= -\frac{1}{2}\pi_{\mathfrak{h}}[\pi_{\mathfrak{p}} \theta_G \wedge \pi_{\mathfrak{p}}\theta_G].
    \end{aligned}
  \]
	\begin{part}{(c)}
		What is the meaning of the torsion of this connection? Compute it.
	\end{part}

	The torsion of a connection on a principal bundle associated to a reductive homogenous space $G/H$ is the obstruction to horizontals being integrable.
	Let $\omega = \pi_\mathfrak{p} \theta_G\in \Omega^1(G;\mathfrak{p})$ be the solder form. Note that it is horizontal and $H$-equivariant descends to a form in $\Omega^1(G/H;\mathfrak{p})$. Recall that the torsion is given by
	\[
		\tau = d\omega +[\Theta\wedge \omega] = d(\pi_{\mathfrak{p}}\theta_G) + [\pi_{\mathfrak{h}}\theta_G \wedge \pi_{\mathfrak{p}}\theta_G].
	\]
	By the Muarer-Cartan equations, this simplifies to
	\[
		\tau = -\frac{1}{2}\pi_{\mathfrak{p}}[\pi_\mathfrak{p}\theta_G, \pi_{\mathfrak{p}} \theta_G].
	\]
	In particular, symmetric spaces admit a canonical torsion-free connection.

	\begin{part}{(d)}
		What is the meaning of geodesics of this connection? Compute them.
	\end{part}

	Any curve $\gamma : \R \to G/H$ can be lifted to a curve $\widetilde{\gamma} : \R \to G$. Geodesic curves must satisfy $\pi_{\mathfrak{h}}\theta_G(\widetilde{\gamma}'(t))=0$. The only explicit solutions to this are $\widetilde{\gamma}(t)=\exp(tX)$ for some $X\in \mathfrak{p}$. Thus every geodesic of the canonical connection on $G/H$ must be of the form $\gamma(t) = \exp(tX)H$ for some $X\in \mathfrak{p}$.


	\begin{part}{(e)}
		Consider the transitive action of $G\times G$ on $G$ by left and right multiplication. Is this homogeneous space reductive?
	\end{part}

	The action can be defined as $(g,h)\cdot q = gqh^{-1}$. Note that the stabilizer of the identity is the diagonal $\Delta\subset G\times G$ with Lie algebra the diagonal $\Delta_\mathfrak{g}\subset \mathfrak{g}\oplus\mathfrak{g}$. The complement can be given by
	\[
		\mathfrak{p}=\left\{ (\xi, -\xi) : \xi \in \mathfrak{g}\right\}\subset\mathfrak{g}\oplus\mathfrak{g}.
	\]
	It should be clear that $\mathfrak{p}$ is an invariant complement so the homogeneous space is indeed reductive.
\end{parts}
%
% \begin{problem}{6}
% \end{problem}
% \begin{parts}
%   \begin{part}{(a)}
%     Let $(\Sigma, g)$ be a Riemannian $2$-manifold, and suppose $\phi : \Sigma \to \R$ is a smooth function. Compute the Gauss curvature $K'$ of the metric $e^{2\phi}g$ in terms of $\phi$ and the Gauss curvature $K$ of the metric $g$.
%   \end{part}
% \end{parts}

\begin{problem}{7}
	Let $G$ be a Lie group with a left-invariant Riemannian metric, denoted $\langle -,-\rangle$. Suppose $\xi,\eta,\nu,\zeta$ are left-invariant vector fields on $G$.
\end{problem}

\begin{parts}
	\begin{part}{(a)}
		Prove that
		\[
		\nabla_\xi \eta = \frac{1}{2}([\xi, \eta] - \ad_\xi^*\eta - \ad_\eta^*\xi),
		\]
		where the star denotes the adjoint with respect to the inner product on the Lie algebra.
	\end{part}

	Recall that the Levi-Civita covariant derivative satisfies the Koszul formula, which here states
	\[
		2\langle \nabla_\xi \eta, \nu\rangle = 
		\langle [\xi, \eta], \nu\rangle - 
		\langle [\xi, \nu], \eta\rangle -
		\langle [\eta, \nu], \xi\rangle
	\]
	since all of the terms of the form $\xi\langle \eta,\nu\rangle$ vanish due to left-invariance. Recall that the adjoint operator behaves as $\ad_X(Y)=[X,Y]$, so we can write
	\[
		\begin{aligned}
		2\langle \nabla_\xi \eta, \nu\rangle 
		&= \langle [\xi, \eta], \nu\rangle - 
		\langle \ad_\xi \nu, \eta\rangle -
	\langle \ad_\eta \nu, \xi\rangle\\
	2\langle \nabla_\xi \eta,\nu\rangle
		&= \langle [\xi,\eta], \nu\rangle - 
		\langle  \nu, \ad_\xi^* \eta\rangle -
	\langle \nu, \ad_\eta^* \xi\rangle\\
	\langle\nabla_\xi \eta, \nu\rangle
		&= \frac{1}{2}\left(\langle [\xi,\eta], \nu\rangle - 
		\langle  \nu, \ad_\xi^* \eta\rangle -
	\langle \nu, \ad_\eta^* \xi\rangle\right)\\
	\nabla_\xi \eta
		&= \frac{1}{2}\left([\xi,\eta] - 
		\ad_\xi^* \eta-
	 \ad_\eta^* \xi\right)\\
		\end{aligned}
\]

\begin{part}{(b)}
	Derive a formula for $\langle R(\xi, \eta)\nu, \zeta\rangle$, where $R$ is the Riemann curvature tensor.
\end{part}
Similar tedious application of formulas as in the last part gives
\[
	\begin{aligned}
	\langle R(\xi,\eta)\nu, \zeta\rangle=\frac{1}{4}\left( 
		\langle [\xi,\eta], [\nu,\zeta]\rangle
		-\langle [\xi,\nu], [\eta,\zeta]\rangle
		-\langle [\eta,\nu], [\xi,\zeta]\rangle
\right)\\
+ \langle B(\xi, \eta), B(\nu,\zeta)\rangle
- \langle B(\xi,\nu), B(\eta,\zeta)\rangle
- \langle B(\eta,\nu), B(\xi, \zeta)\rangle,
\end{aligned}
\]
where $B(X,Y) = (\ad_X^* Y + \ad_Y^* X)/2$.

\begin{part}{(c)}
	Derive a formula for $\langle R(\xi, \eta)\eta, \xi\rangle$, which is the sectional curvature.
\end{part}

The previous problem then gives us
\[
	\begin{aligned}
	\langle R(\xi,\eta)\eta, \xi\rangle=\frac{1}{4}\left( 
		\langle [\xi,\eta], [\eta,\xi]\rangle
		-\langle [\xi,\eta], [\eta,\xi]\rangle
\right)
+ \langle B(\xi, \eta), B(\eta,\xi)\rangle
- \langle B(\xi,\eta), B(\eta,\xi)\rangle = 0.
\end{aligned}
\]

\begin{part}{(d)}
	How do your formulas simplify for bi-invariant Riemannian metrics?
\end{part}

In the first formula, we have $\ad_\xi^* = -\ad_xi$ and so $B=0$. Thus, the formula simplifies to
\[
	\langle R(\xi,\eta)\nu, \zeta\rangle=\frac{1}{4}\left( 
		\langle [\xi,\eta], [\nu,\zeta]\rangle
		-\langle [\xi,\nu], [\eta,\zeta]\rangle
		-\langle [\eta,\nu], [\xi,\zeta]\rangle
\right).
\]
\end{parts}

\end{document}
